---
title: "Reproducible Research Course Project 1"
output: 
        html_document:
                keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data



```{r}
activityData <- read.csv("./activity.csv")
summary(activityData)

```

## What is mean total number of steps taken per day?

First we calculate the total number of steps taken per day
```{r}
totalStepsByDay <- aggregate(steps~date, activityData, sum)

```

Make histogram of the total number of steps taken each day

```{r}
hist(totalStepsByDay$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "green", ylim = c(0, 20), breaks = seq(0,25000, by=2500))

```

Mean of the total number of steps taken per day

```{r}
mean(totalStepsByDay$steps)

```

Median of the total number of steps taken per day

```{r}
median(totalStepsByDay$steps)

```

## What is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
averageStepsByInterval <- aggregate(steps~interval, activityData, mean)
names(averageStepsByInterval) <- c("interval", "mean")
plot(averageStepsByInterval$interval, averageStepsByInterval$mean, type = "l", col = "green", lwd = 2, xlab = "Interval", ylab = "Average number of steps", main = "Average number of steps per interval")

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
averageStepsByInterval[which.max(averageStepsByInterval$mean),]$interval
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
sum(is.na(activityData$steps))
```

Fill in the missing values in the dataset by using the mean/median for that day, or the mean for that 5-minute interval, etc.

First we get the values to be filled up

```{r}
imputed_steps <- averageStepsByInterval$mean[match(activityData$interval, averageStepsByInterval$interval)]
```

Create a new dataset that is equal to the original dataset but with the missing data filled in

```{r}
activityImputed <- transform(activityData, steps = ifelse(is.na(activityData$steps), yes = imputed_steps, no = activityData$steps))

totalStepsImputed <- aggregate(steps ~ date, activityImputed, sum)

names(totalStepsImputed) <- c("date", "daily_steps")
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 

```{r}
hist(totalStepsImputed$daily_steps, col = "green", xlab = "Total steps taken per day", ylim = c(0, 30), main = "Total number of taken per day", breaks = seq(0,25000, by = 2500))

```

Mean of the total number of steps taken per day

```{r}
mean(totalStepsImputed$daily_steps)

```

Median of the total number of steps taken per day

```{r}
median(totalStepsImputed$daily_steps)

```


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activityImputed$date <- as.Date(strptime(activityImputed$date, format="%Y-%m-%d"))
activityImputed$datetype <- sapply(activityImputed$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) == "Sunday")
            {y <- "Weekend"} else
            {y <- "Weekday"}
             y
     
        })

activityDate <- aggregate(steps ~ interval + datetype , activityImputed, mean)
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
library(ggplot2)
plot <- ggplot(activityDate, aes(x = interval, y = steps, color = datetype)) +
        geom_line() +
        labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
        facet_wrap(~datetype, ncol = 1, nrow = 2)
        
print(plot)        
```






